---
title: "Functionalizing rasters from NEON"
author: "Nate Looker"
date: "June 21, 2016"
output: html_document
---

## Objectives

1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)
1. For the CHM, set values == 0 to NA (not trees)
1. Classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that <2m is generally grasses / understory, <6m small trees,and the rest are tall trees. A function could import the desired thresholds. 
1. Visualize histogram/density and plot vertical cutoff lines.
1. Take the chm and create a hillshade (http://neon-workwithdata.github.io/neon-data-institute-2016/R/create-hillshade-R/)
1. PLOT - layer the classified raster on top of the hillshade, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)
1. Export the plot figure to a pdf – publishable
1. Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.


## Import spatial libraries

```{r get_lib}

library(raster)
library(rgdal)
library(cowplot)

```

## Import canopy height model, exclude zero values, and reclassify

```{r read-in-chm}
f_dir <- "../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif"
chm_raw <- raster(f_dir)
# plot density function with vertical lines indicating specified quantiles
dens_w_breaks <- function(data, probs) {
  data[data==0] <- NA
  breaks <- unname(quantile(data, probs))
  density(data, main="Canopy heights at Teakettle site\n with bin cutoffs in red")
  abline(v=breaks, col = "red")
}
# Try out function with breaks at 0, 0.333, and 0.667
perc <- seq(0,1,length=4)
dens_w_breaks(chm_raw, perc)
pdf(file = "TEAK_CHM_density_with_breaks.pdf", width = 6, height = 7)
dens_w_breaks(chm_raw, perc)
dev.off()
# Read in file and reclassify by specified quantiles
bin_CHM <- function(file, probs) {
  chm <- raster(file)
  chm[chm==0] <- NA
  breaks <- unname(quantile(chm, probs))
  chm <- cut(chm, breaks=breaks)
  return(chm)
}


chm <- bin_CHM(f_dir, perc)
breaks <- unname(quantile(chm_raw, perc))

pdf(file = "Plot_of_recl_CHM.pdf", width = 6, height = 5)
par(xpd = F, mar = c(5.1, 4.1, 4.1, 4.5))
plot(chm, col = c("cyan", "orange", "red"), main = "Canopy height model reclassified by elevation quantiles", legend = F)

# Enable plotting legend outside bounds
par(xpd = T)

# Legend x position
leg.x <- par()$usr[2] + 20

# Legend x position
leg.y <- par()$usr[4] + 50 - abs(par()$usr[3] - par()$usr[4])/2

# add the legend
legend(leg.x, leg.y,
       legend = c("Lowest", "Intermediate", "Highest"), fill = c("cyan", "orange", "red"))

dev.off()
```

## Save reclassified raster to GeoTIFF

```{r write-raster}

writeRaster(chm, filename = "reclassified_CHM_TEAK.tif", 
            format = "GTiff",
            options = "COMPRESS=LZW",
            overwrite = T,
            NAflag = -9999)

```


