---
title: "Functionalizing rasters from NEON"
author: "Nate Looker"
date: "June 21, 2016"
output: html_document
---

## Objectives

1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)
1. For the CHM, set values == 0 to NA (not trees)
1. Classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that <2m is generally grasses / understory, <6m small trees,and the rest are tall trees. A function could import the desired thresholds. 
1. Visualize histogram/density and plot vertical cutoff lines.
1. Take the chm and create a hillshade (http://neon-workwithdata.github.io/neon-data-institute-2016/R/create-hillshade-R/)
1. PLOT - layer the classified raster on top of the hillshade, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)
1. Export the plot figure to a pdf – publishable
1. Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.


## Import spatial libraries

```{r get_lib}

library(raster)
library(rgdal)
library(cowplot)

```

## Import canopy height model, exclude zero values, and reclassify

```{r read-in-chm}
f_dir <- "../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif"
temp_chm <- raster(f_dir)
# plot density function with vertical lines indicating specified quantiles
dens_w_breaks <- function(data, probs) {
  data[data==0] <- NA
  breaks <- unname(quantile(data, probs))
  density(data)
  abline(v=breaks)
}
# Try out function with breaks at quintiles
perc_20 <- seq(0,1,by=0.2)
dens_w_breaks(temp_chm, perc_20)


# Read in file and reclassify by specified quantiles
bin_CHM <- function(file, probs) {
  chm <- raster(file)
  chm[chm==0] <- NA
  breaks <- unname(quantile(chm, probs))
  chm <- cut(chm, breaks=breaks)
  return(chm)
}


chm <- bin_CHM(f_dir, seq(0,1,length= 4))
plot(chm, main="Canopy height model reclassified by elevation quantiles")
```

## Plot density

